---
title: "FinalAssignment"
author: "JYOTSNA"
date: "2022-10-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part A - Data Science Questions (12 marks)
There are four questions in this part, with differing marks. All answers must be recorded in an Rmarkdown file then exported to PDF/HTML file.

(3 marks) From your understanding of ethical data science, mention three principles of a code of ethics that any data scientist should consider.

Write your answer as:
P1:___________________________________________________________________
P2:___________________________________________________________________
P3:___________________________________________________________________
(3 marks) To build a visualisation using the ggplot2 library, we use the following template:

ggplot(data= [dataset], mapping = aes(x = [x-variable], y = [y-variable]))+
  geom_xxx() +
  other options 
Based on the above template, mention the main components of building a graph using ggplot2 and describe the meaning of each of these components.

Write your answer as:
______________________________________________________________________
______________________________________________________________________
______________________________________________________________________
______________________________________________________________________
______________________________________________________________________
(3 marks) Describe three properties of the correlation coefficient of two variables

Write your answer as:
1.  ___________________________________________________________________
2.  ___________________________________________________________________
3.  ___________________________________________________________________
(3 marks) Imagine we have a dataset that lists the heights of the fathers and their sons. You have built a linear model that encodes the relationship between the fathers’ heights and the sons’ heights as follows:

lm(son ~ father, data = heights_data)

Call:
lm(formula = son ~ father, data = heights_data)

Coefficients:
(Intercept)    father  
70.45       0.50  
The estimated coefficient (i.e. intercept and slope), which describes the relationship between the fathers’ and sons’ heights can be interpreted as:

Write your answer as:
_____________________________________________________________________
_____________________________________________________________________
_____________________________________________________________________
_____________________________________________________________________
_____________________________________________________________________


# --------------------------------------------------------------------------------- #

# Part B: Data Preparation, exploring and modelling (78 marks)
Data Description:
The four CSV files are described in the following table:
  -Covid19.csv: Master file which includes information on continents,countries, and the daily new cases and the daily new deaths in each country
  -Tests.csv:information about daily covid19 tests in eah country
  -Countries.csv:information about the countries
  -Recovered.csv:information about the daily recovered cases in each country

### Task 1: Data Preparation and Wrangling: (21 marks)
Load and read the data from the CSV files and store them into dataframes named appropriately.

Tidy up the dataframe driven from the file “Recovered.csv” to be compatible with the dataframe driven from the file “Covid19.csv”, i.e., every observation should have a record of recovered patients in one country in a single day.

Change the column names in the dataframes were loaded from the following files accordingly.



```{r}

```

